---
title: "Extroversion Data Analysis"
subtitle: "Model 3 Results"
author: "R. Noah Padgett"
date: "`r Sys.Date()`"
output: html_document
---

```{r set-up, message=FALSE, error=FALSE, warning=FALSE}

# Load packages & utility functions
source("code/load_packages.R")
source("code/load_utility_functions.R")
# environment options
options(scipen = 999, digits=3)

```

```{r}

# Load diffIRT package with data
library(diffIRT)
data("extraversion")
mydata <- na.omit(extraversion)


```


# Model 3: IFA with RT only

## Model details

```{r model3-code}

cat(read_file(paste0(w.d, "/code/JAGS/model_3.txt")))

```


## Model results

```{r model3, cache=TRUE, autodep=TRUE}

# Save parameters
jags.params <- c("tau", "lambda", "theta", "reli.omega", "reli.maximal", "lambda.std", "gamma")
# initial-values
jags.inits <- function(){
    list(
      "tau"=matrix(c(-0.64, -0.09, -1.05, -1.42, -0.11, -1.29, -1.59, -1.81, -0.93, -1.11),
                   ncol=1, nrow=10),
      "lambda"=rep(0.7,10),
      "eta"=rnorm(142),
      "ystar"=matrix(c(0.7*rep(rnorm(142),10)), ncol=10)
    )
  }
jags.data <- list(
  y = mydata[,1:10],
  lrt = mydata[,11:20],
  N = nrow(mydata),
  nit = 10,
  ncat = 2
)
model.fit <-  R2jags::jags(
  model = paste0(w.d, "/code/JAGS/model_3.txt"),
  parameters.to.save = jags.params,
  data = jags.data,
  n.chains = 4,
  n.burnin = 5000,
  n.iter = 10000
)

print(model.fit, width=1000)

```

## Posterior Distribution Summary

```{r}
jags.mcmc <- as.mcmc(model.fit)
a <- colnames(as.data.frame(jags.mcmc[[1]]))
fit.mcmc <- data.frame(as.matrix(jags.mcmc, chains = T, iters = T))
colnames(fit.mcmc) <- c("chain", "iter", a)
fit.mcmc.ggs <- ggmcmc::ggs(jags.mcmc) # for GRB plot

# save posterior draws for later
write.csv(x=fit.mcmc, file=paste0(getwd(),"/data/posterior_draws_m3.csv"))
```

### Categroy Thresholds ($\tau$)

```{r post-summary-tau}
# tau
bayesplot::mcmc_areas(fit.mcmc, regex_pars = "tau", prob = 0.8); ggsave("fig/model3_tau_dens.pdf")
bayesplot::mcmc_acf(fit.mcmc, regex_pars = "tau"); ggsave("fig/model3_tau_acf.pdf")
bayesplot::mcmc_trace(fit.mcmc, regex_pars = "tau"); ggsave("fig/model3_tau_trace.pdf")
ggmcmc::ggs_grb(fit.mcmc.ggs, family = "tau"); ggsave("fig/model3_tau_grb.pdf")
```

### Factor Loadings ($\lambda$)

```{r post-summary-lambda}

bayesplot::mcmc_areas(fit.mcmc, regex_pars = "lambda", prob = 0.8)
bayesplot::mcmc_acf(fit.mcmc, regex_pars = "lambda")
bayesplot::mcmc_trace(fit.mcmc, regex_pars = "lambda")
ggmcmc::ggs_grb(fit.mcmc.ggs, family = "lambda")
```

```{r post-summary-lambda-std}

bayesplot::mcmc_areas(fit.mcmc, regex_pars = "lambda.std", prob = 0.8); ggsave("fig/model3_lambda_dens.pdf")
bayesplot::mcmc_acf(fit.mcmc, regex_pars = "lambda.std"); ggsave("fig/model3_lambda_acf.pdf")
bayesplot::mcmc_trace(fit.mcmc, regex_pars = "lambda.std"); ggsave("fig/model3_lambda_trace.pdf")
ggmcmc::ggs_grb(fit.mcmc.ggs, family = "lambda.std"); ggsave("fig/model3_lambda_grb.pdf")
```

### Latent Response Total Variance ($\theta$)

```{r post-summary-theta}

bayesplot::mcmc_areas(fit.mcmc, regex_pars = "theta", prob = 0.8); ggsave("fig/model3_theta_dens.pdf")
bayesplot::mcmc_acf(fit.mcmc, regex_pars = "theta"); ggsave("fig/model3_theta_acf.pdf")
bayesplot::mcmc_trace(fit.mcmc, regex_pars = "theta"); ggsave("fig/model3_theta_trace.pdf")
ggmcmc::ggs_grb(fit.mcmc.ggs, family = "theta"); ggsave("fig/model3_theta_grb.pdf")
```

### Misclassification Probabilities

```{r post-summary-misclass, cache=TRUE, autodep=TRUE}

A <- fit.mcmc %>%
  dplyr::select(contains("gamma"))

newnames <- colnames(A)
for(i in 1:length(newnames)){
  newnames[i] <- ifelse(
    nchar(newnames[i])==14, 
    substr(newnames[i], 7, 13),
    substr(newnames[i], 7, 14)
  )
  newnames[i] <- str_replace_all(newnames[i], ",","_")
  newnames[i] <- paste0("c_", newnames[i])
}
colnames(A) <- newnames

A <- A %>%
  #select(all_of(pg1)) %>%
  pivot_longer(
    cols = everything(),
    names_to = c("person", "item", "p1", "p2"),
    names_pattern = "c_(.)_(.*)_(.)_(.)",
    values_to = "est"
  )

A %>%
  mutate(para = paste0("p",p1,p2))%>%
  group_by(person, item, para)%>%
  summarise(
    est = mean(est, na.rm=T)
  )%>%
  na.omit()%>%
  ggplot(aes(y=est, x=para)) + 
    geom_boxplot()+
    facet_wrap(.~item)+
    theme_classic()
    
```


### Factor Reliability Omega ($\omega$)

```{r post-summary-omega}
bayesplot::mcmc_areas(fit.mcmc, regex_pars = "reli.omega", prob = 0.8); ggsave("fig/model3_omega_dens.pdf")
bayesplot::mcmc_acf(fit.mcmc, regex_pars = "reli.omega"); ggsave("fig/model3_omega_acf.pdf")
bayesplot::mcmc_trace(fit.mcmc, regex_pars = "reli.omega"); ggsave("fig/model3_omega_trace.pdf")
ggmcmc::ggs_grb(fit.mcmc.ggs, family = "reli.omega"); ggsave("fig/model3_omega_grb.pdf")

# extract omega posterior for results comparison
extracted_omega <- data.frame(model_3 = fit.mcmc$reli.omega)
write.csv(x=extracted_omega, file=paste0(getwd(),"/data/extracted_omega_m3.csv"))

```

### Factor Reliability maximal

```{r post-summary-maximal}

bayesplot::mcmc_areas(fit.mcmc, regex_pars = "reli.maximal", prob = 0.8); ggsave("fig/model3_maximal_dens.pdf")
bayesplot::mcmc_acf(fit.mcmc, regex_pars = "reli.maximal"); ggsave("fig/model3_maximal_acf.pdf")
bayesplot::mcmc_trace(fit.mcmc, regex_pars = "reli.maximal"); ggsave("fig/model3_maximal_trace.pdf")
ggmcmc::ggs_grb(fit.mcmc.ggs, family = "reli.maximal"); ggsave("fig/model3_maximal_grb.pdf")


# extract maximal posterior for results comparison
extracted_omega <- data.frame(model_3 = fit.mcmc$reli.maximal)
write.csv(x=extracted_omega, file=paste0(getwd(),"/data/extracted_maximal_m3.csv"))

```

## Posterior Predictive Distributions

The different posterior predictive distributions are possible for these types of models.
First, using the modeling implied response probability after accounting for misclassification (i.e.g, using the $\omega$) parameters that are a combination of the item factor and response time models.
Secondly, using only the model implied response probabilied from the item factor model.

```{r model3-ppd, cache=TRUE, autodep=TRUE}

# Posterior Predictive Check
Niter <- 200
model.fit$model$recompile()
fit.extra <- rjags::jags.samples(model.fit$model, variable.names = "omega", n.iter = Niter)
N <- 142
nit <- 10
nchain<-4
C <- 2
n <- i <- iter <- ppc.row.i <- 1
y.prob.ppc <- array(dim=c(Niter*nchain, nit, C))
for(chain in 1:nchain){
  for(iter in 1:Niter){
    # initialize simulated y for this iteration
    y <- matrix(nrow=N, ncol=nit)
    # loop over item
    for(i in 1:nit){
      # simulated data for item i for each person
      for(n in 1:N){
        y[n,i] <- sample(1:C, 1, prob = fit.extra$omega[n, i, 1:C, iter, chain])
      }
      # computer proportion of each response category
      for(c in 1:C){
        y.prob.ppc[ppc.row.i,i,c] <- sum(y[,i]==c)/N
      }
    }
    
    # update row of output
    ppc.row.i = ppc.row.i + 1
  }
}

yppcmat <- matrix(c(y.prob.ppc), ncol=1)
z <- expand.grid(1:(Niter*nchain), 1:nit, 1:C)
yppcmat <- data.frame(  iter = z[,1], nit=z[,2], C=z[,3], yppc = yppcmat)

ymat <- model.fit$model$data()[["y"]]
y.prob <- matrix(ncol=C, nrow=nit)
for(i in 1:nit){
  for(c in 1:C){
    y.prob[i,c] <- sum(ymat[,i]==c-1)/N
  }
}
yobsmat <- matrix(c(y.prob), ncol=1)
z <- expand.grid(1:nit, 1:C)
yobsmat <- data.frame(nit=z[,1], C=z[,2], yobs = yobsmat)
plot.ppc <- full_join(yppcmat, yobsmat)

p <- plot.ppc %>%
  mutate(C    = as.factor(C),
         item = nit) %>%
  ggplot()+
  geom_boxplot(aes(x=C,y=y.prob.ppc), outlier.colour = NA)+
  geom_point(aes(x=C,y=yobs), color="red")+
  lims(y=c(0, 1))+
  labs(y="Posterior Predictive Category Proportion", x="Item Category")+
  facet_wrap(.~nit, nrow=1)+
  theme_bw()+
  theme(
    panel.grid = element_blank(),
    strip.background = element_rect(fill="white")
  )
p
ggsave(filename = "fig/model3_ppc_misclass_y.pdf",plot=p,width = 6, height=3,units="in")
ggsave(filename = "fig/model3_ppc_misclass_y.png",plot=p,width = 6, height=3,units="in")
ggsave(filename = "fig/model3_ppc_misclass_y.eps",plot=p,width = 6, height=3,units="in")

```


```{r model3-ppd-pi, cache=TRUE, autodep=TRUE}

# Posterior Predictive Check
Niter <- 200
model.fit$model$recompile()
fit.extra <- rjags::jags.samples(model.fit$model, variable.names = "pi", n.iter = Niter)

N <- 142
nit <- 10
nchain<-4
C <- 2
n <- i <- iter <- ppc.row.i <- 1
y.prob.ppc <- array(dim=c(Niter*nchain, nit, C))
for(chain in 1:nchain){
  for(iter in 1:Niter){
    # initialize simulated y for this iteration
    y <- matrix(nrow=N, ncol=nit)
    # loop over item
    for(i in 1:nit){
      # simulated data for item i for each person
      for(n in 1:N){
        y[n,i] <- sample(1:C, 1, prob = fit.extra$pi[n, i, 1:C, iter, chain])
      }
      # computer proportion of each response category
      for(c in 1:C){
        y.prob.ppc[ppc.row.i,i,c] <- sum(y[,i]==c)/N
      }
    }
    
    # update row of output
    ppc.row.i = ppc.row.i + 1
  }
}

yppcmat <- matrix(c(y.prob.ppc), ncol=1)
z <- expand.grid(1:(Niter*nchain), 1:nit, 1:C)
yppcmat <- data.frame(  iter = z[,1], nit=z[,2], C=z[,3], yppc = yppcmat)

ymat <- model.fit$model$data()[["y"]]
y.prob <- matrix(ncol=C, nrow=nit)
for(i in 1:nit){
  for(c in 1:C){
    y.prob[i,c] <- sum(ymat[,i]==c-1)/N
  }
}
yobsmat <- matrix(c(y.prob), ncol=1)
z <- expand.grid(1:nit, 1:C)
yobsmat <- data.frame(nit=z[,1], C=z[,2], yobs = yobsmat)
plot.ppc <- full_join(yppcmat, yobsmat)

p <- plot.ppc %>%
  mutate(C    = as.factor(C),
         item = nit) %>%
  ggplot()+
  geom_boxplot(aes(x=C,y=y.prob.ppc), outlier.colour = NA)+
  geom_point(aes(x=C,y=yobs), color="red")+
  lims(y=c(0, 1))+
  labs(y="Posterior Predictive Category Proportion", x="Item Category")+
  facet_wrap(.~nit, nrow=1)+
  theme_bw()+
  theme(
    panel.grid = element_blank(),
    strip.background = element_rect(fill="white")
  )
p
ggsave(filename = "fig/model3_ppc_pi_y.pdf",plot=p,width = 6, height=3,units="in")
ggsave(filename = "fig/model3_ppc_pi_y.png",plot=p,width = 6, height=3,units="in")
ggsave(filename = "fig/model3_ppc_pi_y.eps",plot=p,width = 6, height=3,units="in")

```


